### 2017-03-24 autocontrol_review/response_letter.tex

%\documentclass[12pt,draftcls,onecolumn]{IEEEtran}
%\documentclass[12pt,twocolumn]{IEEEtran}
\documentclass[twocolumn]{IEEEtran}
\usepackage{times}

% Packages, commands --------------------------------------------------------------------
\input{pack}

% BIBLIOGRAHY -----------------------------------------------------------------------------
% Option 1: bibtex natbib
%\usepackage[numbers]{natbib}
%\usepackage[colorlinks=true,linkcolor=blue,citecolor=purple,pagebackref=true]{hyperref}
% Option 2: biblatex 
%\usepackage[backend=bibtex,style=ieee-alphabetic,natbib=true,backref=false]{biblatex}
\usepackage[backend=bibtex,style=ieee,natbib=true,backref=false]{biblatex}
\addbibresource{ref.bib}
% Adding back references to bib entries:
\DefineBibliographyStrings{english}{%
  backrefpage = {page},% originally "cited on page"
  backrefpages = {pages},% originally "cited on pages"
}
% Title, authors, etc. -------------------------------------------------------------------------

%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09
\title{A Linearly Relaxed Approximate Linear Program for Markov Decision Processes}
%\author{Chandrashekar Lakshminarayanan$^\star$, Shalabh Bhatnagar$^\star$,
 and Csaba Szepesvari$^\dagger$\thanks{$^\star$Department of Computer
Science and Automation, Indian Institute of Science, Bangalore 560012.
E-mail: $\{$chandrul, shalabh$\}$@csa.iisc.ernet.in}
\thanks{$^\dagger$Department of Computing Science, University of Alberta,
Edmonton, Alberta, Canada T6G 2E8. E-mail: csaba.szepesvari@ualberta.ca}}
% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

% Doc start ------------------------------------------------------------------------------------
\begin{document}
%\maketitle
Dear Editor,
            We would like to re-submit the revised version of our paper. We now list the differences between the previous and current submission.
\begin{itemize}
\item The title of the paper has been changed to ``A Linearly Relaxed Approximate Linear Program for Markov Decisions Processes''.
\item The length of current draft has been increased to 8 pages.
\item We have incorporated major revisions in the paper and restructure the presentation.
\item The length of the introduction has been increased to make the motivation clear. Proper placement with respect to existing literature has been done.
\item The main result has been stated up front alongside a discussion.
\item Streamlined formulation of the main result, with additional corollary that illustrates the utility. Added examples.
\item The proof has been simplified and restructured. The constants in the bound have been slightly improved.
\item Numerical illustration: New example has been added to illustrate the choices available when choosing constraints.
\end{itemize}
We have taken care of all the reviewer comments, except for the following comment by Reviewer~$1$.
\emph{Comment}: Explain the intuition behind stationary Markov deterministic policies being optimal.
\emph{Our Response:} Markov property buys this. Since we are limited by space constraints we have added reference to literature.
\end{itemize}
\end{document}

### 2017-03-24 supplementary/plt.tex
response-letter-single-col-supplement

\FloatBarrier
\begin{figure}[htp]
\begin{minipage}{0.5\textwidth}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{cc}
\begin{tikzpicture}[scale=1,font=\Large,]
    \begin{axis}[]
    \addplot[smooth,very thick, mark=diamond, each nth point=100] plot file {./lralp_exp/V_LP};
    \addplot[dashed,very thick, mark=+, each nth point=100] plot file {./lralp_exp/V_pol_CS};
    \addplot[dotted,very thick,mark=square, each nth point=100] plot file {./lralp_exp/V_pol_cone};
    \addlegendentry{$J^*$};
    \addlegendentry{$J_{CS}$};
    \addlegendentry{$J_{LRA}$};

@@

&
\begin{tikzpicture}[scale=1,font=\Large]
    \begin{axis}[]
\addplot[very thick] plot file {./lralp_exp/policy};
\addplot[dashed,very thick] plot file {./lralp_exp/policy_CS};
\addplot[dotted,very thick] plot file {./lralp_exp/policy_cone};
\addlegendentry{$u^*$};
\addlegendentry{$u_{CS}$};
\addlegendentry{$u_{LRA}$};

>>>>

\FloatBarrier
\begin{figure}[htp]
\begin{minipage}{0.5\textwidth}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{cc}
\begin{tikzpicture}[scale=1,font=\Large,]
    \begin{axis}[]
    \addplot[smooth,very thick, mark=diamond, each nth point=100] plot file {../lralp_exp/V_LP};
    \addplot[dashed,very thick, mark=+, each nth point=100] plot file {../lralp_exp/V_pol_CS};
    \addplot[dotted,very thick,mark=square, each nth point=100] plot file {../lralp_exp/V_pol_cone};
    \addlegendentry{$J^*$};
    \addlegendentry{$J_{CS}$};
    \addlegendentry{$J_{LRA}$};

@@

&
\begin{tikzpicture}[scale=1,font=\Large]
    \begin{axis}[]
\addplot[very thick] plot file {../lralp_exp/policy};
\addplot[dashed,very thick] plot file {../lralp_exp/policy_CS};
\addplot[dotted,very thick] plot file {../lralp_exp/policy_cone};
\addlegendentry{$u^*$};
\addlegendentry{$u_{CS}$};
\addlegendentry{$u_{LRA}$};

### 2017-03-24 supplementary/results.tex
response-letter-single-col-supplement

The bound is given in terms of the approximation error of $J^*$ with the basis functions $\Phi= (\phi_1,\dots,\phi_k)$, as well as the deviation between two functions, $\Jalpo,\Jlro: \S \to \Re$, which we define next. In particular,
\begin{align*}
\Jalpo(s) & = \min\{ r^\top\phi(s) \,|\, \Phi r \ge J^*, \, r\in \Re^k \}\,,\\
\Jlro(s)    & = \min\{ r^\top\phi(s) \,|\, W^\top E \Phi r \ge W^\top E J^*, \, r\in \Re^k \}\,,
\end{align*}
where $s\in \S$. Recall that $E: \Re^S \to \Re^{SA}$ is defined so that $(E J)^\top = (J^\top, \dots, J^\top)$, i.e., $E$ stacks its argument $A$-fold. Hence, $W^\top E = \sum_a W_a^\top$. Our strong assumption is that $\Jlro$ is finite-valued. Note that $\Jalpo\ge J^*$ reflects the error due to using the basis functions $(\phi_j)_j$, and the magnitude of the deviation $\Jlro-\Jalpo$ reflects the error introduced due to the relaxed constraint system. \todoc{I commented out the figure. We can discuss whether we want to include it.}
%The function $\Jalpo$ is an upper approximation to $J^*$

@@

Just like \citet{ALP}, we will also require that $\psi$ is a \emph{stochastic Lyapunov-function} for the MDP.  In particular, we require that the $\alpha$-discounted stability coefficient
\begin{align} \label{eq:betadef}
\beta_\psi \doteq \alpha  \max_{a} \norm{P_a \psi}_{\infty,\psi}
\end{align}
is strictly less than one.
This can be seen to imply that $H: (\Re^S,\norm{\cdot}_{\infty,\psi}) \to (\Re^{SA}, \norm{\cdot}_{\infty,\psi})$ is a contraction,
where for $J = (J_1^\top,\dots,J_A^\top)^\top \in \Re^{SA}$ we let $\norm{J}_{\infty,\psi} = \max_a\norm{J}_{\infty,\psi}$.
That $H$ is a contraction will play a crucial role in our results.
Note that the condition $\beta_\psi<1$ is closely related to the condition that for any policy $u$, 
$P_u \psi \le \psi$, which can be viewed as a stability condition on the MDP 
and which appeared in a slightly altered form in studying the stability of MDPs with infinite 
state spaces \citep[e.g.,][]{chemey99a}.
Note that one can always choose $\psi = \one$: $\beta_\one = \alpha<1$. 
With this, we are ready to state our main result:

@@

From linear programming theory, it follows that primal boundedness is equivalent to dual feasibility. 
Since the dual of $\min\{c^\top x\,:\, Ax \ge b \}$ is 
$\max\{ y^\top b\,:\, y \ge 0, c = A^\top y\}$, we get
that a necessary and sufficient condition for $\Jlro$ 
to be finite-valued is that for any $s\in \S$, $\phi(s)$ lies in the conic span, 
$\{ U \lambda\,:\, \lambda\in \Re_+^{SA}\}$, of (the columns) of $U = \Phi^\top E^\top W$.
%(Note that the boundedness of LRALP only requires that $\Phi^\top c$ lies in the conic span of $U$.)
When $W$ is such that its constituents $W_1,\dots,W_A$ are all identical, 
the conic span of $U$ is equal to the conic span of $\Phi^\top W_1$. 
It is particularly instructive 
to consider the case when the common matrix $W_a = (w_{1,a},\dots,w_{m,a})$ ``selects'' the $m$ states,
i.e., when $\{w_{1,a},\dots,w_{m,a}\}= \{ e_{s}\,:\, s\in \S_0 \}$ for some $\S_0\subset \S$, $|\S_0|\le m$,
where $e_s\in \{0,1\}^\S$ are the $s\in \S$ vectors in the standard Euclidean basis.
In this case, the condition that $\phi(s)$ lies in the conic span of $U$ 
is equivalent to that $\phi(s)$ lies in the conic span of $\phi(\S_0) \doteq \{\phi(s')\,:\, s'\in \S_0\}$. 
Thus, to ensure boundedness of $\Jlro$, 
the chosen states should be selected to ``conicly cover'' all the vectors in $\phi(\S)\subset \Re^k$.%
\footnote{The same implies that, under the same condition on $W$, boundedness of the LRALP 
holds if and only if $\sum_s c(s) \phi(s)$ is in the conic span of $\phi(\S_0)$. Note that this is easy to fulfill
if the support of $c$ has a small cardinality by add all states in the support of $c$ to $\S_0$.
}
\todoc{Add reference to the intro to Mausam's book.}

The next theorem shows that magnitudes of the coefficients used in the conic cover control the size of 
$\norm{\Jalpo-\Jlro}_{\infty,\psi}$. 
For the theorem we let $\Lambda\in \Re_+^{\S\times \S_0}$ be the matrix of conic coefficients: 
For any $s\in \S$, $\phi(s) = \sum_{s'\in \S_0} \Lambda(s,s') \phi(s')$. After the theorem we give constructions 
for creating conic covers.
\begin{theorem}
\label{conetheorm}
Assume that $W_1 = \dots = W_A$, $\{w_{1,a}\,dots,w_{m,a}\} = \{ e_s\,:\, s\in \S_0 \}$ 
and that $\phi(\S)$ lies in the conic span of $\phi(\S_0)$ with conic coefficients given by $\Lambda$.
Let $\epsilon = \inf_{r}\norm{ J^* -  \Phi r }_{\infty,\psi}$.
Then, 
\begin{align*}
\norm{\Jalpo-\Jlro}_{\infty,\psi}
\le 
\norm{\Jalpo-J^*}_{\infty,\psi} + (1+ \norm{\Lambda \psi}_{\psi,\infty}) \epsilon\,.
\end{align*}
\end{theorem}
\begin{proof}
Let $r^*$ be such that $\norm{J^* - \Phi r^*}_{\infty,\psi} = \epsilon$ (this exists by continuity) and let 
$\delta = J^* - \Phi r^*$.
Pick any $s\in \S$ and let $r_s = \argmin\{ r^\top \phi(s) \,:\, W^\top E \Phi r \ge W^\top E J^*,\, r\in \Re^k \}$ 
so that $\Jlro(s) = r_s^\top \phi(s)$.
Note that by assumption, for any $s'\in \S_0$, $\Jlro(s') = r_s^\top \phi(s') \ge J^*(s')$.
Now, notice that by definition, $\Jlro \le \Jalpo$ (the LP defining $\Jlro$ is the relaxation of the LP defining $\Jalpo$).
Hence,
$%\begin{align*}
0\le \Jalpo(s) - \Jlro(s) = \Jalpo(s) - J^*(s) + J^*(s) - \Jlro(s)
$% \end{align*}
and $\Jlro(s) = r_s^\top \phi(s) = r_s^\top \sum_{s'\in \S_0} \Lambda(s,s') \phi(s') = \sum_{s'\in \S_0} \Lambda(s,s') \Jlro(s')
\ge \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')$.
Combining this with the previous inequality we get
$%\begin{align*}
0\le
%&\le 
 \frac{\Jalpo(s) - \Jlro(s)}{\psi(s)}\\
 \le
%&\le
\frac{\Jalpo(s)-J^*(s)}{\psi(s)} + \frac{J^*(s) - \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')}{\psi(s)}
%\,.
$. % \end{align*}
Plugging in $J^*(s) = \phi(s)^\top r^*+\delta(s)$, using again that $\phi(s) = \sum_{s'\in \S_0} \Lambda(s,s') \phi(s')$,
and also using the triangle inequality after taking absolute values, we get
\begin{align*}
\MoveEqLeft \frac{|J^*(s) - \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')|}{\psi(s)} \\
& \le
  \frac{|\delta(s)|}{\psi(s)} + \frac{\sum_{s'\in \S_0} \Lambda(s,s') |\delta(s')|}{\psi(s)} \\
%& \le
%  \frac{|\delta(s)|}{\psi(s)} + \frac{1}{\psi(s)} \, \sum_{s'\in \S_0} \Lambda(s,s') \psi(s') \frac{|\delta(s')|}{\psi(s')} \\
& \le
  \frac{|\delta(s)|}{\psi(s)} + \norm{\delta}_{\infty,\psi} \frac{\sum_{s'\in \S_0} \Lambda(s,s') \psi(s')}{\psi(s)} \, .
\end{align*}
Combining this with the previous display and noting that $\norm{\delta}_{\infty,\psi} = \epsilon$ finishes the proof.
\if0
Putting things together, we get 
\begin{align*}
\norm{\Jalpo - \Jlro}_{\infty,\psi}
\le
\norm{\Jalpo(s)-J^*(s)}_{\infty,\psi}
+ 
\norm{\delta}_{\infty,\psi} (1+ \norm{\Lambda \psi}_{\infty,\psi})\,,
\end{align*}
finishing the proof.
\fi
\end{proof}

>>>>

The bound is given in terms of the approximation error of $J^*$ with the basis functions $\Phi= (\phi_1,\dots,\phi_k)$, as well as the deviation between two functions, $\Jalpo,\Jlro: \S \to \Re$, which we define next. In particular,
\begin{align*}
\Jalpo(s) & = \min\{ r^\top\phi(s) \,|\, \Phi r \ge J^*, \, r\in \Re^k \}\,,\\
\Jlro(s)    & = \min\{ r^\top\phi(s) \,|\, W^\top E \Phi r \ge W^\top E J^*, \, r\in \Re^k \}\,,
\end{align*}
where $s\in \S$. Recall that $E: \Re^S \to \Re^{SA}$ is defined so that $(E J)^\top = (J^\top, \dots, J^\top)$, i.e., $E$ stacks its argument $A$-fold. Hence, $W^\top E = \sum_a W_a^\top$. Our strong assumption is that $\Jlro$ is finite-valued. Note that $\Jalpo\ge J^*$ reflects the error due to using the basis functions $(\phi_j)_j$, and the magnitude of the deviation $\Jlro-\Jalpo$ reflects the error introduced due to the relaxed constraint system. 
%The function $\Jalpo$ is an upper approximation to $J^*$

@@

Just like \citet{ALP}, we will also require that $\psi$ is a \emph{stochastic Lyapunov-function} for the MDP.  In particular, we require that the $\alpha$-discounted stability coefficient
\begin{align} \label{eq:betadef}
\beta_\psi \doteq \alpha  \max_{a} \norm{P_a \psi}_{\infty,\psi}
\end{align}
is strictly less than one.
This can be seen to imply that $H: (\Re^S,\norm{\cdot}_{\infty,\psi}) \to (\Re^{SA}, \norm{\cdot}_{\infty,\psi})$ is a contraction,
where for $J = (J_1^\top,\dots,J_A^\top)^\top \in \Re^{SA}$ we let $\norm{J}_{\infty,\psi} = \max_a\norm{J_a}_{\infty,\psi}$.
That $H$ is a contraction will play a crucial role in our results.
Note that the condition $\beta_\psi<1$ is closely related to the condition that for any policy $u$, 
$P_u \psi \le \psi$, which can be viewed as a stability condition on the MDP 
and which appeared in a slightly altered form in studying the stability of MDPs with infinite 
state spaces \citep[e.g.,][]{chemey99a}.
Note that one can always choose $\psi = \one$, which gives $\beta_\one = \alpha<1$. 
With this, we are ready to state our main result:

@@

From linear programming theory, it follows that primal boundedness is equivalent to dual feasibility. 
Since the dual of $\min\{c^\top x\,:\, Ax \ge b \}$ is 
$\max\{ y^\top b\,:\, y \ge 0, c = A^\top y\}$, we get
that a necessary and sufficient condition for $\Jlro$ 
to be finite-valued is that for any $s\in \S$, $\phi(s)$ lies in the conic span, 
$\{ U \lambda\,:\, \lambda\in \Re_+^{SA}\}$, of (the columns) of $U = \Phi^\top E^\top W$.
%(Note that the boundedness of LRALP only requires that $\Phi^\top c$ lies in the conic span of $U$.)
When $W$ is such that its constituents $W_1,\dots,W_A$ are all identical, 
the conic span of $U$ is equal to the conic span of $\Phi^\top W_1$. 
It is particularly instructive 
to consider the case when the common matrix $W_a = (w_{1,a},\dots,w_{m,a})$ ``selects'' the $m$ states,
i.e., when $\{w_{1,a},\dots,w_{m,a}\}= \{ e_{s}\,:\, s\in \S_0 \}$ for some $\S_0\subset \S$, $|\S_0|\le m$,
where $e_s\in \{0,1\}^\S$ are the $s\in \S$ vectors in the standard Euclidean basis.
In this case, the condition that $\phi(s)$ lies in the conic span of $U$ 
is equivalent to that $\phi(s)$ lies in the conic span of $\phi(\S_0) \doteq \{\phi(s')\,:\, s'\in \S_0\}$. 
Thus, to ensure boundedness of $\Jlro$, 
the chosen states should be selected to ``conicly cover'' all the vectors in $\phi(\S)\subset \Re^k$.%
\footnote{The same implies that, under the same condition on $W$, boundedness of the LRALP 
holds if and only if $\sum_s c(s) \phi(s)$ is in the conic span of $\phi(\S_0)$. Note that this is easy to fulfill
if the support of $c$ has a small cardinality by add all states in the support of $c$ to $\S_0$.
}

The next theorem shows that magnitudes of the coefficients used in the conic cover control the size of 
$\norm{\Jalpo-\Jlro}_{\infty,\psi}$. 
For the theorem we let $\Lambda\in \Re_+^{\S\times \S_0}$ be the matrix of conic coefficients: 
For any $s\in \S$, $\phi(s) = \sum_{s'\in \S_0} \Lambda(s,s') \phi(s')$. After the theorem we give constructions 
for creating conic covers.
\begin{theorem}
\label{conetheorm}
Assume that $W_1 = \dots = W_A$, $\{w_{1,a},\dots,w_{m,a}\} = \{ e_s\,:\, s\in \S_0 \}$ 
and that $\phi(\S)$ lies in the conic span of $\phi(\S_0)$ with conic coefficients given by $\Lambda$.
Let $\epsilon = \inf_{r}\norm{ J^* -  \Phi r }_{\infty,\psi}$.
Then, 
\begin{align*}
\norm{\Jalpo-\Jlro}_{\infty,\psi}
\le 
\norm{\Jalpo-J^*}_{\infty,\psi} + (1+ \norm{\Lambda \psi}_{\psi,\infty}) \epsilon\,.
\end{align*}
\end{theorem}
\begin{proof}
Let $r^*$ be such that $\norm{J^* - \Phi r^*}_{\infty,\psi} = \epsilon$ (this exists by continuity) and let 
$\delta = J^* - \Phi r^*$.
Pick any $s\in \S$ and let $r_s = \argmin\{ r^\top \phi(s) \,:\, W^\top E \Phi r \ge W^\top E J^*,\, r\in \Re^k \}$ 
so that $\Jlro(s) = r_s^\top \phi(s)$.
Note that by assumption, for any $s'\in \S_0$, $\Jlro(s') = r_s^\top \phi(s') \ge J^*(s')$.
Now, notice that by definition, $\Jlro \le \Jalpo$ (the LP defining $\Jlro$ is the relaxation of the LP defining $\Jalpo$).
Hence,
$%\begin{align*}
0\le \Jalpo(s) - \Jlro(s) = \Jalpo(s) - J^*(s) + J^*(s) - \Jlro(s)
$% \end{align*}
and $\Jlro(s) = r_s^\top \phi(s) = r_s^\top \sum_{s'\in \S_0} \Lambda(s,s') \phi(s') = \sum_{s'\in \S_0} \Lambda(s,s') \Jlro(s')
\ge \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')$.
Combining this with the previous inequality we get
$%\begin{align*}
0\le
%&\le 
 \frac{\Jalpo(s) - \Jlro(s)}{\psi(s)}\\
 \le
%&\le
\frac{\Jalpo(s)-J^*(s)}{\psi(s)} + \frac{J^*(s) - \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')}{\psi(s)}
%\,.
$. % \end{align*}
Plugging in $J^*(s) = \phi(s)^\top r^*+\delta(s)$, using again that $\phi(s) = \sum_{s'\in \S_0} \Lambda(s,s') \phi(s')$,
and also using the triangle inequality after taking absolute values, we get
\begin{align*}
\MoveEqLeft \frac{|J^*(s) - \sum_{s'\in \S_0} \Lambda(s,s') J^*(s')|}{\psi(s)} \\
& \le
  \frac{|\delta(s)|}{\psi(s)} + \frac{\sum_{s'\in \S_0} \Lambda(s,s') |\delta(s')|}{\psi(s)} \\
%& \le
%  \frac{|\delta(s)|}{\psi(s)} + \frac{1}{\psi(s)} \, \sum_{s'\in \S_0} \Lambda(s,s') \psi(s') \frac{|\delta(s')|}{\psi(s')} \\
& \le
  \frac{|\delta(s)|}{\psi(s)} + \norm{\delta}_{\infty,\psi} \frac{\sum_{s'\in \S_0} \Lambda(s,s') \psi(s')}{\psi(s)} \, .
\end{align*}
Combining this with the previous display and noting that $\norm{\delta}_{\infty,\psi} = \epsilon$ finishes the proof.
\if0
Putting things together, we get 
\begin{align*}
\norm{\Jalpo - \Jlro}_{\infty,\psi}
\le
\norm{\Jalpo(s)-J^*(s)}_{\infty,\psi}
+ 
\norm{\delta}_{\infty,\psi} (1+ \norm{\Lambda \psi}_{\infty,\psi})\,,
\end{align*}
finishing the proof.
\fi
\end{proof}
